### 2020-04-23 paper.tex
Debugging Concepts

\begin{itemize}
	\item Imperative
	\item Declarative
	\item Omniscient
\end{itemize}

>>>>

Debugging concepts can be divided roughly into three categories: Traditional \bf{imperative-focused} debuggers provide basic functionality to interact with programs at runtime: Once the code execution is paused (i.e. using a breakpoint at a given statement), we gain access to the current call stack and the values assigned to variables of a given frame. Manual control of the programs execution allows to inspect the programs behavior step by step as well as assigning new values to variables ``on-the-fly''.

Where individual transformations of an RP based program can be inspected using a traditional debugger, they fall short in understanding the greater model of the dependency graph. In fact, RP provides its own challenges to debuggers. So are call stacks spoiled with the invocation of RP runtime system internal matter or breakpoints can only be used on the imperative parts of transformations rather than any given point in the dependency graph. Various \bf{reactive} debuggers and visualization tools for different RP runtimes emerged over the past few years \cite{REDebuggerSalvaneschi} \cite{RxViz} \cite{RxFiddle} exploring this field. They have knowledge of the dependency graph and therefore allow improved interaction with RP concepts.

In matters of time, traditional as well as reactive debuggers work with the very present state of the programs execution only. They do not have any knowledge about what happened before or what is going to happen in the future. This shortcoming is tedious, especially when debugging a problem depending from multiple complex circumstances in a system. An \bf{omniscient} debugger \cite{omniscientDebugger} does not interact with the executed program directly. Instead, it records runtime telemetry and provides an interface for later inspection. This way, engineers can ``time travel'' back and fore through program execution without the hassle of reconstructing a given failure situation. This concept comes with its own set of challenges, i.e. demands on storage capacity or user experience \emph{(UX)} design. \cite{omniscientDebuggerProblems}
